---
title: "CSCI 385 - Third Deliverable"
author: "Garrett Welton"
date: "12/7/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(RSocrata)
library(caret)
library(ggplot2) 
library(tidyverse)
```

## Improvement

```{r}
#API call
df <- read.socrata(
  "https://chronicdata.cdc.gov/resource/fwns-azgu.csv",
  app_token = "OJxxBUd1kEfoUDA5ACIVnD88u")

#Data Tidying
NHIS <- filter(df, !is.na(data_value), data_value_type == "Age-Standardized" | break_out_category == "Age") %>%
  select(year, locationdesc, category, topic, data_value, break_out_category, break_out) %>%
  rename(location = locationdesc, percent_value = data_value)

#Spit into separate tables
cardioDiseases <- filter(NHIS, category == "Cardiovascular Diseases") %>%
  select(-category) %>%
  pivot_wider(names_from = c(break_out_category, break_out),
              values_from = percent_value) %>%
  select(-`Age_25-44`, -`Age_18-24`, -Race_Other, -`Race_Non-Hispanic Asian`) %>%
  drop_na()
  
cardioDiseases.majorCardiovascularDisease <- filter(cardioDiseases, topic == "Major Cardiovascular Disease")
cardioDiseases.heartAttack <- filter(cardioDiseases, topic == "Acute Myocardial Infarction (Heart Attack)")
cardioDiseases.coronaryHeartDisease <- filter(cardioDiseases, topic == "Coronary Heart Disease")
cardioDiseases.stroke <- filter(cardioDiseases, topic == "Stroke")

riskFactors <- filter(NHIS, category != "Cardiovascular Diseases") %>%
  select(-category) %>%
  pivot_wider(names_from = c(break_out_category, break_out),
              values_from = percent_value,
              values_fn = mean) %>%
  select(-Race_Other, -`Race_Non-Hispanic Asian`) %>%
  drop_na()

riskFactors.physicalInactivity <- filter(riskFactors, topic == "Physical Inactivity")
riskFactors.smoking <- filter(riskFactors, topic == "Smoking")
riskFactors.hypertension <- filter(riskFactors, topic == "Hypertension")

#Data Exploration
ggplot(cardioDiseases.majorCardiovascularDisease, aes(x = year, y = Overall_Overall, color = location))+
  geom_point()+
  geom_smooth(se = FALSE)
```

This graph is a breakdown of the percent of the population of different regions of the US who have Major Cardiovascular Disease over time. It can be used to get an idea of the trends of Major Cardiovascular Disease in theses regions and how they compare.

## Tables

```{r}
Heart <- read.csv("heart.csv")
head(Heart)

head(cardioDiseases.majorCardiovascularDisease)
head(cardioDiseases.heartAttack)
head(cardioDiseases.coronaryHeartDisease)
head(cardioDiseases.stroke)

head(riskFactors.physicalInactivity)
head(riskFactors.smoking)
head(riskFactors.hypertension)
```

## Validation

Five fold cross validation on knn model

```{r}
heart <- as_tibble(Heart) %>% select(-trestbps, -ca)
heart$target = as.factor(heart$target)

set.seed(132)

trControl <- trainControl(method  = "cv",
                          number  = 5)

fit <- train(target ~ .,
             method     = "knn",
             tuneGrid   = expand.grid(k = 3:15),
             trControl  = trControl,
             data       = heart)

fit

#Accuracy plot
ggplot(fit$results, aes(x=k, y=Accuracy)) +
  geom_point() +
  geom_line()
```

This knn model has an accuracy of about 79%, which is over 70%, making it a successful model according to my definition of "Success" and "Failure". It is fairly accurate at predicting the presence of heart disease in patients baised on thier records.

## Implications

The implications of these insights is that a doctor may be able to predict heart disease in patients earlier from patient records. Hopefully helping to reduce related deaths. One ethical issue that may arise is that the model has the potential to be used health insurance companies to increase the rate of or deny coverage to those who who it deems to have heart disease.
